The dynamics of starvation and recovery

### 2017-06-12 Manuscript/NatureComm/YKR_NatureComm.tex
cover letter

\noindent {\bf Energetic barriers to body size}
The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}a).
Similar predictions have been made for carnivore populations using alternative consumer-resource models \citep{DeLong:2012kw}.
Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998}.
The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\citep{Damuth:1987kr,allen2002,enquist1998}. %Previous studies have focused on t
This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\citep{allen2002,enquist1998}.
This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \citep{Damuth:1987kr,allen2002,enquist1998}.
Figure \ref{fig:mass}a shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly constant over this same range.
This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.

@@

\noindent{\bf A within-lineage driver of Cope's Rule}
The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig. \ref{fig:mass}a, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \citep{tilman1981,dutkiewicz2009,barton2010}. Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but these absolute limits do not offer a within-lineage mechanism by which larger body sizes are selected for or against.
We will now show that the NSM indeed provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time known as Cope's Rule.

>>>>

% \noindent {\bf Energetic barriers to body size}
The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}a).
Similar predictions have been made for carnivore populations using alternative consumer-resource models \citep{DeLong:2012kw}.
Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998}.
The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\citep{Damuth:1987kr,allen2002,enquist1998}. %Previous studies have focused on t
This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\citep{allen2002,enquist1998}.
This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \citep{Damuth:1987kr,allen2002,enquist1998}.
Figure \ref{fig:mass}a shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly constant over this same range.
This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.

@@

% \noindent{\bf A within-lineage driver of Cope's Rule}
The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig. \ref{fig:mass}a, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \citep{tilman1981,dutkiewicz2009,barton2010}. Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but these absolute limits do not offer a within-lineage mechanism by which larger body sizes are selected for or against.
We will now show that the NSM indeed provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time known as Cope's Rule.
